Selective electroporation of the viral envelope via optimal control of transmembrane potential

### 2021-06-10 documents/background.tex
Transmembrane library
Pretty graphs

\begin{fquote}[Professor David Baltimore][\cite{Issues2008}]
	The scientific literature is a conversation among scientists; we describe to each other what we have done, and we implicitly ask our colleagues to try to build on our results, testing their value and their ability to support a growing edifice.

@@

Vijayalaxmi \cite{Biological2016} \cite{Comprehensive2018} \cite{Funding2019} already offer a number of excellent guidelines on best practices in this field; as do Chou \cite{chou1996radio} and \cite{Effects2016} . We only offer a few issues which we were particularly concerned about when designing our experiments and searching literature.

@@

>>>>

\begin{fquote}[Professor David Baltimore (recently deceased)][\cite{Issues2008}]
	The scientific literature is a conversation among scientists; we describe to each other what we have done, and we implicitly ask our colleagues to try to build on our results, testing their value and their ability to support a growing edifice.

@@

Vijayalaxmi \cite{Biological2016} \cite{Comprehensive2018} \cite{Funding2019} already offer a number of excellent guidelines on best practices in this field; as do Chou \cite{chou1996radio} and Napotnik \cite{Effects2016} . We only offer a few issues which we were particularly concerned about when designing our experiments and searching literature.

@@

In 1984, O. P. Gandhi \cite{BIOLOGICALa} performed Raman spectroscopy of a variety of viruses in the frequency range 1 to 5e13 Hz.

### 2021-06-10 documents/pure_electroporation.tex
Pretty graphs

\title{Optimal control for selective electroporation of the viral envelope}
\author{\footnotesize{Daniel Correia\orcidlink{0000-0002-9353-0216}}\footnote{Ontario, Canada; contact information in the ORCiD.}}
\date{\small{March 21 2021}}

@@

\begin{abstract}
	\noindent The phenomenon of electroporation occurs when extreme applied electric fields in the MV/m range drive transmembrane potentials, leading to pores in, and perhaps the destruction of, membranes; this is commonly harnessed for transfection into the cell in the laboratory, and has recently been shown clinically to be safe and effective for the in-situ ablation of tumors.

	\noindent As pulse duration is reduced from the microsecond to nanosecond regime, effects targeting smaller intracellular structures are expected and observed, and a degree of tunability arises. Indeed, pulses shorter than 10 ns have been observed to porate the nuclear envelope and organelle membranes while leaving the plasma membrane intact.

@@

when a strong electric field is applied, they "do enough things to justify international meetings, to fill a sizable book, and to lead one to speak of an entirely new technology for cell manipulation."

@@

Despite considering the somewhat harder problem of electroporating while inside the cell, rather than in extracellular fluid, \cite{Electroporation2013} report that 250 nm liposomes should be porated without inducing damage to the host cell, and modelling shows that 100 nm liposomes may be porated by 4 ns pulses never exceeding 20 MV/m.

The same field intensities do not appear to damage phages with protein capsids\cite{Manipulation2013}, in line with our negative results on T4.

Non-resonant capsid and envelope breakage with 120 pulses, 500 ns each, at 3 MV/m, has been previously observed\cite{Inactivation1990}, but the RNA damage suggests that this is an electrochemical artifact \cite{Formation1996} (c.f. \cite{Microwave1987}).

@@

Viruses have extraordinary mechanical properties\footnote{Why is this? What selection pressures drive the selection of this pressure?}, so it seems reasonable that their dielectric properties could differ substantially from those of other biological structures.

In the extreme case of the bacteriophage, the genome is packaged so tightly into the protein capsid that the core is often described as a glassy crystal of DNA\cite{Conformation2007}. ssRNA viruses like coronaviridae might be expected to be somewhat more sedate in packing density, but this does not seem to be borne out in the core conductivity data.

On the other hand, the eukaryotic "structome" also contains countless vesicles and structures of similar nature that might plausibly be damaged; small pockets of lipid in the endoplasmic reticula - nuclear pore complexes, etc. This appears to be seen to some degree in Gabriel's data; dense tissues exhibit many [double check this]. It may not be possible to meaningfully predict all possible off-target effects, much like in drug discovery.

@@

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host porates at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m. Mutatoyo's data also agree.

@@

@@

Despite the two having nearly identical time constants, the course of the transmembrane voltage is clearly different.
[transmembrane plot here]

@@

Kotnik note an important fact: all these permittivity and conductivity values will also depend on frequency, which is not included in their model. We have not examined how these factors can be extracted from. Also disregarded are the Poisson-boltzmann shielding space-charges that proved all-important in the last study.

@@

The concept of optimizing an electroporation pulse is not novel, having been discussed by many groups in the past; either to prodduce different sterilization effects, tee Retelj use a series of, to

>>>>

\title{Selective electroporation of the viral envelope via optimal control of transmembrane potential}
\author{\footnotesize{Daniel Correia\orcidlink{0000-0002-9353-0216}}\footnote{Ontario, Canada; contact information in the ORCiD.}}
\date{\small{March 21 2021}}

@@

\begin{abstract}
	\noindent The phenomenon of electroporation occurs when extreme applied electric fields in the MV/m range drive transmembrane potentials, leading to pores in (and potentially the destruction of) membranes. This is commonly harnessed for transfection into the cell in the laboratory, and has recently been shown clinically to be safe and effective for the in-situ ablation of tumors.

	\noindent As pulse duration is reduced from the microsecond to sub-nanosecond regime, effects targeting smaller intracellular structures are expected and observed, and a degree of tunability arises. Indeed, pulses shorter than 10 ns have been observed to porate the nuclear envelope and organelle membranes while leaving the plasma membrane intact.

@@

when a strong electric field is applied, they "...do enough things to justify international meetings, to fill a sizable book, and to lead one to speak of an entirely new technology for cell manipulation."

@@

Despite considering the presumably harder problem of electroporating small lipid vesicles while inside the cell, rather than in extracellular fluid, \cite{Electroporation2013} report that 250 nm liposomes should be porated without inducing damage to the host cell, and their modelling shows that 100 nm liposomes may be porated by 4 ns pulses never exceeding 20 MV/m.

The same field intensities do not appear to damage phages with protein capsids\cite{Manipulation2013}, in line with our previously-reported negative results on T4.

Capsid and envelope permeablization with 120 pulses, 500 ns each, at 3 MV/m, has been previously observed\cite{Inactivation1990}, but the RNA damage suggests that this is an electrochemical artifact \cite{Formation1996} (c.f. \cite{Microwave1987}).

@@

Viruses have extraordinary mechanical properties (why is this? What selection pressures drive the selection of this pressure?), so it seems reasonable to speculate that their dielectric properties could differ substantially from those of other biological structures.

In the extreme case of coliform bacteriophage, the genome is packaged so tightly into the protein capsid that the core is most accurately described as a glassy crystal of DNA\cite{Conformation2007}. ssRNA viruses like coronaviridae might be expected to be somewhat more sedate in packing density, but this does not seem to be borne out in the core conductivity data.

On the other hand, the eukaryotic "structome" also contains countless vesicles and structures of similar nature that might plausibly be damaged; small pockets of lipid in the endoplasmic reticula - nuclear pore complexes, etc. This appears to be seen to some degree in Gabriel's data; dense tissues exhibit a high cole-cole alpha coefficient [double check this]. It may not be possible to meaningfully predict all possible off-target effects, much like in drug discovery.

@@

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host becomes permeable at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m. Mutatoyo's data also agree.

@@

A normalized (note that the true value of the virus' transmembrane voltage is, again, a factor of 400 reduced).

\begin{figure}[H]
	\includegraphics[width=\textwidth]{step_course}
\end{figure}

Despite the two having similar time constants (host: $\tau_1=444\ \text{ns},\ \tau_2=2.3\ \text{ns}$, virus: $\tau_1=295\ \text{ns},\ \tau_2=3.1\ \text{ns}$), the time course of the transmembrane voltage is clearly distinct. (Plot generated via the convolution method, \ghfile{transmembrane_steady_state.py})

Consider the following double gaussian waveform, each with a FWHM of 500 picoseconds, separated by 1 nanosecond. In this case, the peak virus transmembrane voltage is no longer 1/400 less than the host cell, but 1/6.5, and the total energy in each curve is also improved to a ratio of 1/37. The second, inverted pulse has the effect of quickly zeroing out the host cell's decaying potential to improve the virus' energy ratio.

\begin{figure}[H]
	\includegraphics[width=\textwidth]{gaussian_pulse}
\end{figure}

@@

@@

Kotnik note an important fact: all these permittivity and conductivity values will also depend on frequency, which is not included in their model. We have not examined whether these factors can be extracted from existing plots. Also disregarded are the Poisson-Boltzmann shielding space-charges that proved all-important in the last study.

@@

The concept of optimizing an electroporation pulse is not novel, having been discussed by many groups in the past; either to produce different sterilization effects, tee Retelj use a series of, to
